Details matter: Similar-looking host galaxies can produce different merging compact binary populations

### 2015-05-26 paper/paper.tex
paper.tex: More editing

\newcommand\mc{{\cal M}_c{}}
\newcommand\editremark[1]{{\color{red}#1}}
\newcommand\unit[1]{\text{#1}}
\newcommand\abbrvPSgrb{PSG}
\newcommand\abbrvPSellipticals{PSE}
\newcommand\abbrvKBLowZa{BD2010}
\newcommand\ForInternalReference[1]{}
\newcommand\modelDefault{A}
\newcommand\modelNiino{A'}
\newcommand\modelKW{D}
\newcommand\modelMannucci{D'}
\newcommand\modelKB{B}
\newcommand\modelPanter{C}
\newcommand\cqg{CQG}
%\newcommand\aap{A\&A}
%\newcommand\apss{APSS}
%\newcommand\aaps{AAPS}
%% \newcommand\apjs{ApJ S}
%% \newcommand\aj{AJ}
%% \newcommand\apjl{ApJL}
%% \newcommand\mnras{MNRAS}
%% \newcommand\pasp{PASP}
%% \newcommand\pasj{PASJ}
%% \newcommand\araa{ARAA}
%% \newcommand\physrep{Phys. Rep.}

@@

The simulation h277 (Boring Galaxy) is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{Loebman12,Loebman14,Kassin14}, baryon fraction
\citep{Munshi13}, halo properties \citep{Zolotov09,Zolotov10), and
  disk properties \citep{Brooks11}.  It has a virial mass of $M_{vir}
  = 6.79 \times 10^{11}$ \msun, stellar mass $M_* = 4.24 \times
  10^{10}$ \msun, and circular velocity $v_{circ} = xx$ km s$^{-1}$.

  The simulation h258 (Exciting Galaxy), on the other hand, has a much
  more active merger history.  At $z = 1$ there is a 1:1 merger event;
  a gas disk rapidly reforms following the collision (see
  \citet{Governato09}), resulting in a massive disk galaxy at $z = 0$
  which looks remarkably similar to the Milky Way and to the other
  simulation, h277.  Prior to the $z = 1$ merger, each of the two
  progenitor galaxies actually experiences its own additional major
  merger events around $z = 2.5$ ({\bf check!}).  The combination of
  the series of major mergers, plus a number of minor interactions and
  flybys, gives a stark contrast to the relatively quiescent history
  of h277.  At $z = 0$, h258 has a virial mass of $M_{vir} = 7.74
  \times 10^{11}$ \msun, stellar mass $M_* = 4.46 \times 10^{10}$
  \msun, and circular velocity $v_{circ} = xx$ km s$^{-1}$.

@@

\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
The black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels.  Because of the extremely narrow range of masses that permit tidal disruption in BH-NS binaries, small changes
in the mass distribution can produce significant changes in the electromagnetically detected population.
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

@@

\appendix

>>>>

\newcommand\mc{{\cal M}_c{}}
\newcommand\editremark[1]{{\color{red}#1}}
\newcommand\msun{M_\odot}
\newcommand\unit[1]{\text{#1}}
\newcommand\abbrvPSgrb{PSG}
\newcommand\abbrvPSellipticals{PSE}
\newcommand\abbrvKBLowZa{BD2010}
\newcommand\ForInternalReference[1]{}
\newcommand\modelDefault{A}
\newcommand\modelNiino{A'}
\newcommand\modelKW{D}
\newcommand\modelMannucci{D'}
\newcommand\modelKB{B}
\newcommand\modelPanter{C}
\newcommand\cqg{CQG}
%\newcommand\aap{A\&A}
%\newcommand\apss{APSS}
%\newcommand\aaps{AAPS}
%% \newcommand\apjs{ApJ S}
%% \newcommand\aj{AJ}
%% \newcommand\apjl{ApJL}
%% \newcommand\mnras{MNRAS}
%% \newcommand\pasp{PASP}
%% \newcommand\pasj{PASJ}
%% \newcommand\araa{ARAA}
%% \newcommand\physrep{Phys. Rep.}

@@

The simulation h277 (Boring Galaxy) is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{Loebman12,Loebman14,Kassin14}, baryon fraction
\citep{Munshi13}, halo properties \citep{Zolotov09,Zolotov10}, and
  disk properties \citep{Brooks11}.  It has a virial mass of $M_{vir}
  = 6.79 \times 10^{11}  \msun$, stellar mass $M_* = 4.24 \times
  10^{10} \msun$, and circular velocity $v_{circ} = \editremark{xx}$ km s$^{-1}$.

  The simulation h258 (Exciting Galaxy), on the other hand, has a much
  more active merger history.  At $z = 1$ there is a 1:1 merger event;
  a gas disk rapidly reforms following the collision (see
  \citet{Governato09}), resulting in a massive disk galaxy at $z = 0$
  which looks remarkably similar to the Milky Way and to the other
  simulation, h277.  Prior to the $z = 1$ merger, each of the two
  progenitor galaxies actually experiences its own additional major
  merger events around $z = 2.5$ ({\bf check!}).  The combination of
  the series of major mergers, plus a number of minor interactions and
  flybys, gives a stark contrast to the relatively quiescent history
  of h277.  At $z = 0$, h258 has a virial mass of $M_{vir} = 7.74
  \times 10^{11} \msun$, stellar mass $M_* = 4.46 \times 10^{10}$
  \msun, and circular velocity $v_{circ} = \editremark{xx}$ km s$^{-1}$.

@@

\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
A neutron star  (NS) coalescing with a black hole (BH) will be tidally disrupted prior to merger if and only if the component
masses, spins, and equation of state permit it.  As described below, only a small fraction of all  possible coalescing
BH-NS binaries have parameters consistent with tidal disruption.  
%
Stellar evolution and hence supernovae remnants are expected to evolve
substantially with birth metallicity, the black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels and the fraction of coalescing NS-BH binaries which are tidally disrupted will vary
over cosmic time. 
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

\begin{figure}
\includegraphics{Figures/fig-mma-TidalDisrupt-MbhA-VersusTheta}
\caption{\label{fig:TidalDisrupt}\textbf{Tidal disruption}:  Following  \citet{2012PhRvD..86l4007F}, a plot of
  the critical spin needed for tidal disruption versus black hole mass and inclination (solid black, dotted black,
  dashed black, solid blue, and dotted blue are $\theta_{LS}=\pi/2,3\pi/8,\pi/4, \pi/8,0$ respectively).  Specifically,
  this figure corresponds to $M_{\rm disk}=0$ in Eq. (\ref{eq:Mdisk}), assuming  a $1.4 M_\odot$ neutron star with compactness ${\cal
    C}=1.66$ in inclined orbit around a spinning black hole of mass $m_{bh}$ and dimensionless spin $\chi_{bh}$.
  Parameters consistent with tidal disruption are above the curves; prompt capture occurs below the
  curve.
}
\end{figure}

% POINT: Trend versus mass and opening angle.  [Be careful: disruption time will depend on opening angle]
To estimate whether a specific compact binary will be tidally disrupted prior to merger, we employ the
following procecure to identify posterior samples consistent with tidal disruption of a neutron star.  First, the
smaller mass must lie within the range of masses allowed by our fiducial NS equation of state (\textbf{MPA1 from Ben}), here
between $0.5 M_\odot$ and $2.5 M_\odot$.  Second, the disruption process must leave behind a remnant disk with nonzero
mass, as estimated using Foucart's expression \cite{2012PhRvD..86l4007F} (his Eqs. (6,12-13)):
\begin{align}
\frac{M_{disk}}{M_{NS}} &=
 0.288 (3 m_{bh}/m_{ns})^{1/3}[1-2 {\cal C}] 
\nonumber \\
&
 -0.148 (m_{bh}/m_{ns})  {\cal C} R_{isco}(a)/m_{bh} 
\label{eq:Mdisk}
\end{align}
where $q=m_{bh}/m_{NS}$ is the binary mass ratio; $a=S_{bh}/m_{bh}^2$ is a dimensionless measure of the black hole spin;
and  where ${\cal C}(m_{ns})=m_{ns}/R_{ns}$ is a mass- and equation-of-state dependent  measure of the neutron star
compactness.  In Foucart's expression, shown above, $R_{isco}(a)$ is the radius of the  innermost stable equitorial circular orbit  of a test particle
about a Kerr black hole  \cite{1972ApJ...178..347B}. % Brandon:  [their Eq. 2.21].   
When the black hole spin is not aligned with the orbit, we use the aligned component of the black hole spin magnitude $a\rightarrow a \cos \theta_{LS}$ in this
expression.   Fixing the neutron star mass $m_{ns}=1.4 M_\odot$ and hence compactness to ${\cal C}=1.66$, this expression implies a two-dimensional surface in
the three-dimensional parameter space $m_{bh},a,\theta_{LS}$ seperates parameters corresponding to tidal disruption to
parameters corresponding to prompt coalescence.
%
After coalescence, the remnant disk generates a tightly beamed electromagnetic counterpart.  For simplicity, we assume
beaming angle is independent of the disruption process or progenitor parameters, so a constant beaming correction factor $f_b$ enters into all
subsequent calculations and can hence be omitted.  
%
Figure \ref{fig:TidalDisrupt} shows contours of this expression for different values of $\theta_{LS}$.

\begin{figure}
\caption{\textbf{Disruption fraction}: Approximate tidal disrpution fraction}
\end{figure}

% POINT: Distribution of tilt angles
Based on Monte Carlo simulations of supernova kicks \editremark{citations:Eggleton,Kalogera,IhmXXX}, we approximate the tilt angle distribution $p(\theta)$ of bound,
coalescing BH-NS binaries by the positive half of a normal distribution
\[
p(\theta)\simeq e^{-\theta^2/2\sigma_\theta^2} \frac{2}{\sqrt{2\pi \sigma_\theta^2}}
\]
where $\sigma_{\theta}\simeq 1$ is a characteristic angular width that varies relatively weakly with the black hole mass
and supernova kick velocity, for velocities consistent with pulsar recoil velocities \editremark{citations}.  This simple distribution is consistent with some detailed population synthesis simulations by
\texttt{StarTrack}. \editremark{validate with more modern startrack}  
For simplicity, in this work we adopt $\sigma_{\theta}=1$.   
%  POINT Average over spin distributions
Using this approximate distribution of misalignment angles, Figure  assumed black hole mass and spin we calculate the
fraction of coalescing binaries consistent with

% POINT: Average over distributions

The relative probability of a given tilt angle

- first aproximations: tilt angle independent of merger time

% POINT: Include delay time: get examples of disruption fraction versus time

------

@@

\appendix
\section{Tidal disruption probability}
